---
title: "Exam LBG-FS2022 - Problem 2: Relationship and Inbreeding"
output: html_notebook
---

```{r}
set.seed(1298)
```


## Disclaimer
Documentation and Protocol for engineering data and results for problem 2.


## Package
Try using the package 

```{r}
if (!is.element("pedsuite", installed.packages()))
  install.packages("pedsuite", dependencies = TRUE)
```

Test pedigrees are shown below

### Cousin Pedigree

```{r}
library(pedsuite)
x <- cousinPed(1, child = TRUE)
plot(x)
```

The generated pedigree can be converted to a tibble

```{r}
library(dplyr)
tbl_cousin <- tibble::as_tibble(x)
(tbl_cousin_with_parents <- tbl_cousin %>%
  filter(fid != 0 & mid != 0) %>%
    select(id, fid, mid))

```

Adding ages of owners to each of the animals

```{r}
(tbl_cousin_with_parents_with_owner_age <- tbl_cousin_with_parents %>% 
  bind_cols(tibble::tibble(AgeOwner = round(rnorm(nrow(tbl_cousin_with_parents), mean = 75, sd = 3), digits = 0))))
```

Remap the data to IDs which are from different spaces.

```{r}
vec_ids <- base::union(base::union(tbl_cousin_with_parents$id, tbl_cousin_with_parents$fid), tbl_cousin_with_parents$mid)
tbl_id_map <- tibble::tibble(src_id = vec_ids,
                             trg_id = sapply(runif(length(vec_ids), min = 114000220, max = 114000399), function(x) paste0("CH", floor(x)), USE.NAMES = FALSE))
tbl_id_map
```

```{r}
tbl_cousin_with_parents_with_owner_age_remap <- tbl_cousin_with_parents_with_owner_age %>% 
  left_join(tbl_id_map, by = c("id" = "src_id")) %>%
  rename(Animal = trg_id)
tbl_cousin_with_parents_with_owner_age_remap <- tbl_cousin_with_parents_with_owner_age_remap %>%
  left_join(tbl_id_map, by = c("fid" = "src_id")) %>%
  rename(Sire = trg_id)
tbl_cousin_with_parents_with_owner_age_remap <- tbl_cousin_with_parents_with_owner_age_remap %>%
  left_join(tbl_id_map, by = c("mid" = "src_id")) %>%
  rename(Dam = trg_id)
(tbl_cousin_with_parents_with_owner_age_remap <- tbl_cousin_with_parents_with_owner_age_remap %>%
  select(Animal, Sire, Dam, AgeOwner))

```



Writing the pedigree to a file

```{r}
s_ped1_path <- "lbgfs2022_exam_problem2_pedigree1.csv"
readr::write_delim(tbl_cousin_with_parents_with_owner_age_remap, s_ped1_path, delim = ",")
```



## Half-Sib Pedigree with Inbred Child

```{r}
x1 = halfSibPed(nch1 = 1, nch2 = 2, sex1 = 1, sex2 = 2:1) |> 
  addSon(parents = 4:5)
plot(x1)
```

Converting it to a tibble

```{r}
tbl_half_sib <- tibble::as_tibble(x1)
(tbl_half_sib_with_parents <- tbl_half_sib %>%
  filter(fid != 0 & mid != 0) %>%
    select(id, fid, mid))
```

Adding ages of owners to each of the animals

```{r}
(tbl_half_sib_with_parents_with_owner_age <- tbl_half_sib_with_parents %>% 
  bind_cols(tibble::tibble(AgeOwner = round(rnorm(nrow(tbl_half_sib_with_parents), mean = 54, sd = 4), digits = 0))))
```

Remap the data to IDs which are from different spaces.

```{r}
vec_ids <- base::union(base::union(tbl_half_sib_with_parents$id, tbl_half_sib_with_parents$fid), tbl_half_sib_with_parents$mid)
tbl_id_map <- tibble::tibble(src_id = vec_ids,
                             trg_id = sapply(runif(length(vec_ids), min = 115000220, max = 115000399), function(x) paste0("CH", floor(x)), USE.NAMES = FALSE))
tbl_id_map
```


```{r}
tbl_half_sib_with_parents_with_owner_age_remap <- tbl_half_sib_with_parents_with_owner_age %>% 
  left_join(tbl_id_map, by = c("id" = "src_id")) %>%
  rename(Animal = trg_id)
tbl_half_sib_with_parents_with_owner_age_remap <- tbl_half_sib_with_parents_with_owner_age_remap %>%
  left_join(tbl_id_map, by = c("fid" = "src_id")) %>%
  rename(Sire = trg_id)
tbl_half_sib_with_parents_with_owner_age_remap <- tbl_half_sib_with_parents_with_owner_age_remap %>%
  left_join(tbl_id_map, by = c("mid" = "src_id")) %>%
  rename(Dam = trg_id)
(tbl_half_sib_with_parents_with_owner_age_remap <- tbl_half_sib_with_parents_with_owner_age_remap %>%
  select(Animal, Sire, Dam, AgeOwner))

```

Writing the pedigree to a file

```{r}
s_ped2_path <- "lbgfs2022_exam_problem2_pedigree2.csv"
readr::write_delim(tbl_half_sib_with_parents_with_owner_age_remap, s_ped2_path, delim = ",")
```


## Data
The data is taken from pedigrees of WAS (`/qualstorzws01/data_projekte/projekte/genmon/pedigrees/PopReport_WAS_mit_20210115.csv_adaptfin5.csv`): 


```
1010779957|1004634509|1005998580|2020-10-29|M|3942|0||0
1010779958|1004634509|1005998580|2020-10-29|F|3942|0||0
1005998580|1000171975|1000719353|2016-10-02|F|3942|0||0
1009031358|1004634509|1005998580|2018-09-15|M|3942|0||0
1004634509|999147646|999222110|2015-05-02|M|3984|0||0
```

This data is not needed so far.


## Solution
In subtask a) the numerator relationship matrices must be constructed

### Numerator Relationship Matrix

```{r}
tbl_ped1 <- readr::read_delim(file = s_ped1_path,
                              delim = ",",
                              col_types = readr::cols(
                                Animal = readr::col_character(),
                                Sire = readr::col_character(),
                                Dam = readr::col_character(),
                                AgeOwner = readr::col_double()
                              ))
tbl_ped1
```

Check whether sire and dams are already in pedigree and determine the founder animals to be added to the pedigree

```{r}
vec_sire_founder <- unique(tbl_ped1$Sire)
vec_sire_founder <- vec_sire_founder[!is.element(vec_sire_founder, tbl_ped1$Animal)]
vec_dam_founder <- unique(tbl_ped1$Dam)
vec_dam_founder <- vec_dam_founder[!is.element(vec_dam_founder, tbl_ped1$Animal)]
vec_founder <- c(vec_sire_founder, vec_dam_founder)
```

Add founders to the pedigree

```{r}
n_nr_founder <- length(vec_founder)
tbl_ped1_founder <- tibble::tibble(Animal = vec_founder, Sire = rep(NA, n_nr_founder), Dam = rep(NA, n_nr_founder), AgeOwner = rep(NA, n_nr_founder))
tbl_ped1_aug <- dplyr::bind_rows(tbl_ped1_founder, tbl_ped1)
tbl_ped1_aug
```

To be able to use `pedigreemm` for computing the numerator relationship matrix, animals must have integer IDs. Hence, we map the character IDs back to integer IDs.

```{r}
tbl_ped_ID_map <- tibble::tibble(DomainCharId = tbl_ped1_aug$Animal,
                                 ImageIntId = c(1:length(tbl_ped1_aug$Animal)))
tbl_ped_ID_map
```

Do the mapping via joins

```{r}
library(dplyr)
# Animals
tbl_ped1_int_id <- tbl_ped1_aug %>%
  left_join(tbl_ped_ID_map, by = c("Animal" = "DomainCharId")) %>%
  rename(AnimalIntId = ImageIntId)
# Sire
tbl_ped1_int_id <- tbl_ped1_int_id %>%
  left_join(tbl_ped_ID_map, by = c("Sire" = "DomainCharId")) %>%
  rename(SireIntId = ImageIntId)
# Dam 
tbl_ped1_int_id <- tbl_ped1_int_id %>%
  left_join(tbl_ped_ID_map, by = c("Dam" = "DomainCharId")) %>%
  rename(DamIntId = ImageIntId)
tbl_ped1_int_id
```

The numerator relationship

```{r}
ped_ped1 <- pedigreemm::pedigree(sire = tbl_ped1_int_id$SireIntId, 
                                 dam =  tbl_ped1_int_id$DamIntId,
                                 label = as.character(tbl_ped1_int_id$AnimalIntId))
pedigreemm::getA(ped = ped_ped1)
```

Same for pedigree 2

```{r}
# read data
tbl_ped2 <- readr::read_delim(file = s_ped2_path,
                              delim = ",",
                              col_types = readr::cols(
                                Animal = readr::col_character(),
                                Sire = readr::col_character(),
                                Dam = readr::col_character(),
                                AgeOwner = readr::col_double()
                              ))
tbl_ped2

# add founders
vec_sire_founder <- unique(tbl_ped2$Sire)
vec_sire_founder <- vec_sire_founder[!is.element(vec_sire_founder, tbl_ped2$Animal)]
vec_dam_founder <- unique(tbl_ped2$Dam)
vec_dam_founder <- vec_dam_founder[!is.element(vec_dam_founder, tbl_ped2$Animal)]
vec_founder <- c(vec_sire_founder, vec_dam_founder)
n_nr_founder <- length(vec_founder)
tbl_ped2_founder <- tibble::tibble(Animal = vec_founder, Sire = rep(NA, n_nr_founder), Dam = rep(NA, n_nr_founder), AgeOwner = rep(NA, n_nr_founder))
tbl_ped2_aug <- dplyr::bind_rows(tbl_ped2_founder, tbl_ped2)
tbl_ped2_aug

# mapping to integer IDs
tbl_ped_ID_map <- tibble::tibble(DomainCharId = tbl_ped2_aug$Animal,
                                 ImageIntId = c(1:length(tbl_ped2_aug$Animal)))
tbl_ped_ID_map
library(dplyr)
# Animals
tbl_ped2_int_id <- tbl_ped2_aug %>%
  left_join(tbl_ped_ID_map, by = c("Animal" = "DomainCharId")) %>%
  rename(AnimalIntId = ImageIntId)
# Sire
tbl_ped2_int_id <- tbl_ped2_int_id %>%
  left_join(tbl_ped_ID_map, by = c("Sire" = "DomainCharId")) %>%
  rename(SireIntId = ImageIntId)
# Dam 
tbl_ped2_int_id <- tbl_ped2_int_id %>%
  left_join(tbl_ped_ID_map, by = c("Dam" = "DomainCharId")) %>%
  rename(DamIntId = ImageIntId)
tbl_ped2_int_id

#
ped_ped2 <- pedigreemm::pedigree(sire = tbl_ped2_int_id$SireIntId, 
                                 dam =  tbl_ped2_int_id$DamIntId,
                                 label = as.character(tbl_ped2_int_id$AnimalIntId))
pedigreemm::getA(ped = ped_ped2)
```

