---
title: "LBGFS 2022 - Exam: Problem 4: BLUP-Animal Model"
output: html_notebook
---


## Disclaimer


## Pedigree
The pedigree is constructed using the package `pedtools`. The following examples are taken from the vignette of pedtools (https://cran.r-project.org/web/packages/pedtools/vignettes/pedtools.html)

```{r}
x = pedtools::ancestralPed(g = 2) # bottom person is "7"

# Bottom part
y = pedtools::halfCousinPed(degree = 1) # top person is "2"
y = pedtools::swapSex(y, 4)
#> Changing sex of spouses as well: 5

# Merge
z = pedtools::mergePed(x, y, by = c("7" = "2"), relabel = TRUE)

# plot
pedtools::plotPedList(list(x, y, z))
```

Merging two half cousin pedigrees

```{r}
p1 <- pedtools::halfCousinPed(degree = 1, child = TRUE)
plot(p1)
```

```{r}
p2 <- pedtools::halfCousinPed(degree = 1, removal = 1)
plot(p2)
```

Merge the pedigree

```{r}
p_merged <- pedtools::mergePed(p1, p2, by = c("10" = "2"), relabel = TRUE)
pedtools::swapSex(p_merged, id = "17")
pedtools::swapSex(p_merged, id = "20")
plot(p_merged)
```

As a table, the pedigree is

```{r}
tbl_ped_sim <- tibble::as_tibble(p_merged)
# change sex of 17 and 20
tbl_ped_sim$sex[c(17,20)] <- 3 - tbl_ped_sim$sex[c(17,20)]
tbl_ped_sim
```

Reduce the pedigree for the dataset and rename parent columns

```{r}
library(dplyr)
tbl_data_p04 <- tbl_ped_sim %>%
  filter(fid != 0 & mid != 0) %>%
  select(-sex) %>%
  rename(sire = fid, dam = mid)
tbl_data_p04
```

Add a herd 

```{r}
tbl_data_p04 <- dplyr::bind_cols(tbl_data_p04,
                                 tibble::tibble(herd = c(rep(1L,3), rep(2L,4), rep(3L,3))))
n_nr_obs <- nrow(tbl_data_p04)
tbl_data_p04
```


## Simulate Observations
Get numerator relationship matrix. Start by adding founders

```{r}
vec_founder_sire <- unique(tbl_data_p04$sire[!is.element(tbl_data_p04$sire, tbl_data_p04$id)])
vec_founder_sire
```

```{r}
vec_founder_dam <- unique(tbl_data_p04$dam[!is.element(tbl_data_p04$dam, tbl_data_p04$id)])
vec_founder_dam
```

```{r}
vec_founder_cmb <- c(vec_founder_sire, vec_founder_dam)
```

Add founders

```{r}
n_nr_founder <- length(vec_founder_cmb)
tbl_ped_aug <- dplyr::bind_rows(tibble::tibble(id = vec_founder_cmb, 
                                               sire = rep(NA, n_nr_founder), 
                                               dam = rep(NA, n_nr_founder)),
                                dplyr::select(tbl_data_p04, id, sire, dam))
n_nr_ani_ped <- nrow(tbl_ped_aug)
tbl_ped_aug
```

Numerator relationship matrix

```{r}
ped_p04 <- pedigreemm::pedigree(sire = as.integer(tbl_ped_aug$sire),
                                dam  = as.integer(tbl_ped_aug$dam),
                                label = tbl_ped_aug$id)
mat_A_p04 <- as.matrix(pedigreemm::getA(ped = ped_p04))
mat_A_p04
```

Cholesky factor

```{r}
mat_R_p04 <- t(chol(mat_A_p04))
mat_R_p04
```

### Simulate the data

Fix genetic parameters

```{r}
n_sigma_p2 <- 64
n_h2 <- 0.16
n_sigma_u2 <- n_h2 * n_sigma_p2
n_sigma_e2 <- n_sigma_p2 - n_sigma_u2
n_lambda <- n_sigma_e2 / n_sigma_u2
```

Fixed effects of sex and herd

```{r}
# effects
# vec_sex_eff <- c(2.9, 7.1)
vec_herd_eff <- c(19.7, 5.5, 43.2)
# vec_fix_eff <- c(vec_sex_eff, vec_herd_eff)
vec_fix_eff <- vec_herd_eff
# design matrix
# mat_X <- model.matrix(lm(id ~ 0 + as.factor(sex), data = tbl_data_p04))
# attr(mat_X, "assign") <- NULL
# attr(mat_X, "contrasts") <- NULL
# colnames(mat_X) <- NULL
mat_X <- model.matrix(lm(id ~ 0 + as.factor(herd), data = tbl_data_p04))
attr(mat_X, "assign") <- NULL
attr(mat_X, "contrasts") <- NULL
colnames(mat_X) <- NULL
# mat_X <- cbind(mat_X, mat_X_herd)
mat_X
```

Check rank

```{r}
Matrix::rankMatrix(mat_X)
```

Random animal effect

```{r}
mat_Z <- cbind(matrix(0, nrow = n_nr_obs, ncol = n_nr_founder), diag(nrow = n_nr_obs))
mat_Z
```

Put together observations

```{r}
set.seed(2579)
vec_y_sim <- crossprod(t(mat_X), vec_fix_eff) + 
  crossprod(t(mat_Z), crossprod(t(mat_R_p04), rnorm(n_nr_ani_ped, sd = sqrt(n_sigma_u2)))) + 
  rnorm(n_nr_obs, sd = sqrt(n_sigma_e2))
vec_y_sim
```

Add observations to dataset

```{r}
tbl_data_p04 <- dplyr::bind_cols(tbl_data_p04, tibble::tibble(y = round(vec_y_sim[,1], digits = 2)))
tbl_data_p04
```

Write the data to a file

```{r}
s_data_p04_path <- file.path(here::here(), "docs", "data", "lbgfs2022_exam_problem4.csv")
readr::write_delim(tbl_data_p04, s_data_p04_path, delim = ",")
```


## Analysis
Read the data

```{r}
s_data_root <- "https://charlotte-ngs.github.io/lbgfs2022/data"
s_data_p04_url <- file.path(s_data_root, "lbgfs2022_exam_problem4.csv")
tbl_data_p04_read <- readr::read_delim(s_data_p04_url, 
                                       delim = ",",
                                       col_types = readr::cols(
                                         y = readr::col_double(),
                                         .default = readr::col_integer()
                                       ))
tbl_data_p04_read
```

The inverse numerator relationship matrix

```{r}
ped_p04 <- pedigreemm::pedigree(sire = as.integer(tbl_ped_aug$sire),
                                dam  = as.integer(tbl_ped_aug$dam),
                                label = tbl_ped_aug$id)
mat_A_inv_p04 <- as.matrix(pedigreemm::getAInv(ped = ped_p04))
```

Mixed model equations

```{r}
vec_y_read <- tbl_data_p04_read$y
mat_xtx <- crossprod(mat_X)
mat_xtz <- crossprod(mat_X, mat_Z)
mat_ztx <- crossprod(mat_Z, mat_X)
stopifnot(min(mat_ztx - t(mat_xtz)) == 0)
stopifnot(max(mat_ztx - t(mat_xtz)) == 0)
mat_ztzlAinv <- crossprod(mat_Z) + n_lambda * mat_A_inv_p04
mat_coef <- rbind(cbind(mat_xtx,mat_xtz), cbind(mat_ztx,mat_ztzlAinv))
# rhs
mat_rhs <- rbind(crossprod(mat_X, vec_y_read),
                 crossprod(mat_Z, vec_y_read))
# sol
mat_sol <- solve(mat_coef, mat_rhs)
mat_sol
```

```{r}
nrow(mat_sol)
```


The reliabilities are computed based on PEV

$$\left[ 
\begin{array}{lr}
X^TR^{-1}X  &  X^TR^{-1}Z \\
Z^TR^{-1}X  &  Z^TR^{-1}Z + G^{-1}
\end{array}
\right]
$$

```{r}
n_nr_herd_levels <- ncol(mat_X)
mat_coef_full <- mat_coef / n_sigma_e2
vec_pev <- diag(solve(mat_coef_full))[(n_nr_herd_levels+1):nrow(mat_coef)]
vec_pev
```

```{r}
vec_A_diag <- diag(mat_A_p04)
vec_A_diag
```


```{r}
vec_rel <- 1 - (vec_pev / (n_sigma_u2 * vec_A_diag))
vec_rel
```

```{r}
vec_B_per <- 100 * vec_rel
vec_B_per
```

