---
title: "Estimate Single Locus Breeding Values from Data"
date: "2022-10-07"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
Breeding values are estimated from data


## Read Data
This is the notebook developed in class. It shows how genotype frequencies can be determined using R.

We want to read a dataset with genotype and phenotype information. 

```{r}
s_gen_phen_data <- "https://charlotte-ngs.github.io/lbgfs2022/data/p1_mrk_one_locus.csv"
library(readr)
tbl_gen_phen <- read_delim(file = s_gen_phen_data,
                           delim = ";",
                           col_types = cols(
                             ID = col_integer(),
                             Genotype = col_integer(),
                             Phenotype = col_double()
                           ))
head(tbl_gen_phen)
```


Descriptive statistics

```{r}
library(dplyr)
tbl_gen_phen %>%
  group_by(Genotype) %>%
  summarise(geno_count = n()) %>%
  summarise(geno_freq = geno_count / sum(geno_count))

```


## Genotypic Values
Regression of phenotypes on Genotypes using homozygous genotypes

```{r}
tbl_hom_gen <- tbl_gen_phen %>%
  filter(Genotype < 3)
head(tbl_hom_gen)
```

```{r}
tbl_hom_gen %>%
  group_by(Genotype) %>%
  summarise(geno_count = n())
```

Regression model, with or without intercept? Compute the mean phenotypic value for all animals with Genotype $0$

```{r}
tbl_hom_gen0 <- tbl_hom_gen %>%
  filter(Genotype < 1)
mean(tbl_hom_gen0$Phenotype)
```

Plot the data

```{r}
library(ggplot2)
p_g0 <- ggplot(data = tbl_hom_gen, mapping = aes(x = Genotype, y = Phenotype)) +
  geom_point()
p_g0
```


From the plot and from the mean of the animals with genotype $0$, it becomes clear that we need to fit a regression with intercept

```{r}
lm_geno_val <- lm(Phenotype ~ Genotype, data = tbl_hom_gen)
(sry_geno_val <- summary(lm_geno_val))
```

Add the regression line to the plot

```{r}
p_g0_line <- p_g0 + 
  geom_abline(slope = sry_geno_val$coefficients["Genotype", "Estimate"],
              intercept = sry_geno_val$coefficients["(Intercept)", "Estimate"])
p_g0_line
```

The genotypic value $a$ is

```{r}
(n_geno_a <- sry_geno_val$coefficients["Genotype", "Estimate"])
```

The genotypic values for heterozygous animals

```{r}
tbl_het_gen <- tbl_gen_phen %>%
  filter(Genotype > 2)
head(tbl_het_gen)
```

```{r}
n_geno_d <- mean(tbl_het_gen$Phenotype) - sry_geno_val$coefficients["(Intercept)", "Estimate"] - n_geno_a
n_geno_d
```


```{r}
tbl_het_gen$Genotype <- 1L
head(tbl_het_gen)
```

```{r}
tbl_gen_all <- bind_rows(tbl_hom_gen, tbl_het_gen)
head(tbl_gen_all)
```

```{r}
p_g_all <- ggplot(data = tbl_gen_all, mapping = aes(x=Genotype,  y=Phenotype))+
  geom_point()+ 
  geom_abline(slope = sry_geno_val$coefficients["Genotype", "Estimate"],
              intercept = sry_geno_val$coefficients["(Intercept)", "Estimate"])
p_g_all

```


Now that the genotypic values $a$ and $d$ are computed, we can do the transformations of the phenotypic values such that the three genotypes have the correct genotypic value

```{r}
tbl_trans_gen <- tbl_gen_all
tbl_trans_gen$Phenotype <- tbl_trans_gen$Phenotype - sry_geno_val$coefficients["(Intercept)", "Estimate"] - n_geno_a
```

To check the result of the transformation, we can compute the means of the transformed phenotypic values

```{r}
tbl_trans_gen %>%
  group_by(Genotype) %>%
  summarise(mean_phen = mean(Phenotype))
```
































































