---
title: "Reticulate and Sympy"
date: "2022-10-13"
output:
  pdf_document: default
  html_document: default
---

## Disclaimer
Tests and experiments with `reticulate` and `sympy`


## Basic Tests
From https://rstudio.github.io/reticulate/index.html, we check the basic functionality with the following code chunks


```{r}
library(reticulate)
# use_python(Sys.which("python3"))
```

Simple usage for python

```{python}
print("Hello Python!")
```

Installing pandas, to then be able to read data with pandas

```{r}
# create virtual env
virtualenv_create("r-reticulate")
# install pandas
virtualenv_install("r-reticulate", "pandas")
```

Use virtual env

```{r}
use_virtualenv("r-reticulate")
```


Checking

```{r}
py_available()
```

Module pandas

```{r}
py_module_available("pandas")
```

Now use pandas to import data

```{python}
import pandas
geno = pandas.read_csv("/Users/pvr/Data/Projects/Github/charlotte-ngs/lbgfs2022_gh-root/main/lbgfs2022/docs/data/p1_mrk_one_locus.csv")
print(geno.head())
```


## Installation of 'sympy'
The python package 'sympy' is installed with

```{r, eval=TRUE}
virtualenv_install("r-reticulate", "sympy")
```

Checking whether the module sympy was installed can be done with

```{r}
py_module_available("sympy")
```



## Testing sympy
According to https://cerebralmastication.com/2020/07/27/using-python-sympy-in-a-python-code-chunk-in-rmarkdown/, the following test for sympy should work.

```{python, eval=TRUE, results='asis'}
from sympy import *
x = symbols('x')
a = Integral(cos(x)*exp(x), x)
output = Eq(a, a.doit())
latex_output = latex(output)
print('$$' + latex_output + '$$\n')
```

An alternative way to show the output with an R-code chunk is

```{r}
knitr::asis_output(sprintf("$$%s$$\n", py$latex_output))
```


## Use sympy for Derivations in Quantitative Genetics

```{python, results='asis'}
# import
from sympy import *
# symbol definition
a, d, p, q, mu, mu_res = symbols('a, d, p, q, mu, mu_res')
# allele frequencies
q = 1-p
# compute population mean
mu = a * p**2 + d * 2 * p * q + (-a) * q**2
mu_res = simplify(collect(expand(mu), [a, d]))
print('$$' + latex(mu) + '$$\n')
print('$$' + latex(mu_res) + '$$\n')

```


