---
title: "Breeding Values"
date: "2022-10-07"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Disclaimer
This notebook shows how breeding values for one single locus are computed. 


## Input 
Input parameters consist of the allele frequency $p$ and of the genotypic values $a$ and $d$.

```{r}
p <- 0.1
q <- 1-p
a <- 10
d <- 2
```

Alpha

```{r}
alpha <- a + (q-p) * d
alpha
```

Breeding values

```{r}
BV11 <- 2*q*alpha
BV12 <- (q-p) * alpha
BV22 <- -2*p*alpha
```

Table

```{r}
tbl_bv <- tibble::tibble(Genotype = c("G1G1", "G1G2", "G2G2"),
                         BreedingVAlue = c(BV11, BV12,BV22))
tbl_bv
```


Function for computing breeding values

```{r}
compute_bv <- function(pn_af, pn_geno_a, pn_geno_d){
  n_freq_q <- 1-pn_af
  alpa <- pn_geno_a + (n_freq_q-pn_af)*pn_geno_d
  BV11 <- 2*n_freq_q*alpha
  BV12 <- (n_freq_q-pn_af)*alpha
  BV22 <- -2*pn_af*alpha
  tbl_bv_result <- tibble::tibble(Genotype = c("G1G1", "G1G2", "G2G2"),
                         BreedingVAlue = c(BV11, BV12,BV22))
  return(tbl_bv_result)
}
```

Call the function

```{r}
tbl_bv <- compute_bv(pn_af = 0.1, pn_geno_a = 10, pn_geno_d = 2)
tbl_bv
```

## Difference between Breeding values
The pairwise difference between the breeding values in the above table can be computed as follows.

```{r}
tbl_bv$BreedingVAlue[1] - tbl_bv$BreedingVAlue[2]
```

```{r}
tbl_bv$BreedingVAlue[2] - tbl_bv$BreedingVAlue[3]

```

































