---
title: Livestock Breeding and Genomics - `r tools::toTitleCase(params$doctype)` 4
author: Peter von Rohr
date: 2022-10-14
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: QTL Data
```{r problem1-setup, echo=FALSE}
if (params$isonline){
  s_data_path <- "https://charlotte-ngs.github.io/lbgfs2022/data/p1_qtl_1_loci.csv"  
} else {
  s_data_path <- file.path(here::here(), "docs", "data", "p1_qtl_1_loci.csv")
}
```

Estimate genotypic values $a$ and $d$ and predict breeding values for all animals using the QTL-data given under:

```{r, echo=FALSE, results='asis'}
cat(s_data_path, "\n")
```


<!-- your-solution-start
### Your Solution

* Start by re-formatting the paternal and maternal alleles into a genotype
* Use a linear regression to estimate genotypic values
* Predict breeding values for all animals

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

<!-- master-solution-end -->
 

## Problem 2: 
Change the phenotypic records in the above given dataset such that the QTL explains $50%$ of the genetic variation when a heritability of $0.45$ is assumed. It is assumed that the QTL acts purely additively, hence the genotypic value of the heterozygotes can be set to $d=0$. 

Show the results as a scatter plot of all phenotypic values for the QTL genotypes.

<!-- your-solution-start
### Your Solution


---  your-solution-end -->


<!-- master-solution-start -->

### Solution
In a first step the phenotypic variance, the genetic variance and the QTL-variance must be determined from the data and the given information.

```{r}
tbl_qtl <- readr::read_delim(file = s_data_path, delim = ";")
head(tbl_qtl)
```

<!-- master-solution-end -->
 
